---
title: "Getting Started With Data Analysis"
author: "Adejumo Ridwan Suleiman"
---

## R Data Types

In R, data types represent the different types of values that can be stored in variables or data structures. The following are the commonly used data types in R.

### Numeric Data Type
Numeric data type is used to represent numbers, including integers and floating-point numbers.
```{r}
# Example:
x <- 10
y <- 3.14
```

### Character Data Type
Character data type is used to represent strings of text.
```{r}
# Example:
name <- "John Doe"

```

### Logical Data Type
Logical data type is used to represent logical values - TRUE or FALSE.
```{r}
# Example:
is_raining <- TRUE
is_sunny <- FALSE

```

### Factor Data Type
Factor data type is used to represent categorical variables. It is useful for handling data with a limited number of possible values.
```{r}
# Example:
gender <- factor(c("male", "female", "male", "male", "female"))

```

### Date and Time Data Types
R has built-in data types for representing dates and times.
```{r}
# Example:
date_today <- Sys.Date()  # returns today's date
time_now <- Sys.time()    # returns the current time

```

### Complex Data Type
Complex data type is used to represent complex numbers.

```{r}
# Example:
z <- 3 + 4i

```


### Raw Data Type
Raw data type is used to represent binary data.
```{r}
# Example:
x <- charToRaw("hello world")
```

### Exercise:

Create a vector that stores the ages of a group of people. The ages should be numeric. Create a factor variable that stores the gender of each person. Finally, create a variable that stores the current date and time.


## R Operators

Operators are used to perform operations on values in R. There are several types of operators in R, including arithmetic operators, assignment operators, comparison operators, logical operators, and more.

### Arithmetic Operators
Arithmetic operators are used to perform mathematical operations on numeric values.
```{r}
# Example:
x <- 10
y <- 3
x + y   # addition
x - y   # subtraction
x * y   # multiplication
x / y   # division
x ^ y   # exponentiation
x %% y  # modulus
x %/% y # integer division
```

### Assignment Operators
Assignment operators are used to assign values to variables.
```{r}
# Example:
x <- 10
y <- 3
x <- y   # assign the value of y to x

```

### Comparison Operators
Comparison operators are used to compare values and return logical values (TRUE or FALSE).
```{r}
# Example:
x <- 10
y <- 3
x == y  # equal to
x != y  # not equal to
x > y   # greater than
x < y   # less than
x >= y  # greater than or equal to
x <= y  # less than or equal to

```


### Logical Operators
Logical operators are used to combine logical values and return logical values.
```{r}
# Example:
x <- TRUE
y <- FALSE
x & y   # AND
x | y   # OR
!x      # NOT

```

### Other Operators
Other Operators
R has other operators, including:

- Membership Operators: `%in%` is used to check if an element is a member of a vector.
- Concatenation Operators: `c()` is used to concatenate vectors.
- Ternary Operator: `ifelse()` is used to return values based on a logical condition.
```{r}
# Example:
x <- 1:5
3 %in% x             # returns TRUE
y <- c(6, 7, 8)
z <- c(x, y)         # concatenates x and y
result <- ifelse(x > 3, "high", "low")  # returns "high" if x > 3, "low" otherwise

```

### Exercise:

Write an R code that creates two variables x and y that store any numeric values of your choice. Using the comparison operators, check if x is greater than y and store the result in a variable called "is_greater". Print out the value of "is_greater".

## R Functions

R functions are reusable code blocks that can take input arguments and perform specific tasks. In this article, we will discuss the basics of R functions, how to create them, and how to use them with examples, detailed explanations, and exercises.

### Creating R functions
To create an R function, we use the function keyword followed by the function name, input arguments (if any), and the code block that defines what the function does. Here's an example of a simple R function that adds two numbers:

```{r}
add_two_numbers <- function(x, y) {
  result <- x + y
  return(result)
}
```

In this example, add_two_numbers is the name of the function, and it takes two input arguments x and y. The function code block adds the two input arguments and stores the result in the result variable. The return statement is used to return the result of the function.

### Using R functions
Once we have created an R function, we can use it by calling its name and passing the required input arguments. Here's an example of using the add_two_numbers function we created earlier:

```{r}
sum <- add_two_numbers(5, 10)
print(sum)  # Output: 15
```
In this example, we pass the values 5 and 10 as input arguments to the add_two_numbers function, which returns the sum of the two numbers as 15. We then store this value in the sum variable and print it to the console using the print function.

### Function arguments
R functions can have zero or more input arguments. Input arguments are enclosed in parentheses after the function name, separated by commas. Here's an example of a function that takes no input arguments:
```{r}
hello_world <- function() {
  print("Hello, world!")
}
```

In this example, the hello_world function takes no input arguments and simply prints "Hello, world!" to the console when called.

R functions can also have default values for input arguments. If a default value is provided, the input argument becomes optional, and if the argument is not provided when the function is called, the default value is used. Here's an example of a function with a default value for an input argument:

```{r}
greet_person <- function(name = "John") {
  print(paste("Hello,", name))
}
```

In this example, the greet_person function takes an input argument name, which has a default value of "John". If we call the function without providing a value for name, it will use the default value and print "Hello, John" to the console. If we provide a value for name, the function will use that value instead.

### Function return values
R functions can return a single value or multiple values. To return a value, we use the return statement followed by the value we want to return. If no return statement is used, the function will return the last value calculated in the code block. Here's an example of a function that returns a single value:

```{r}
multiply_two_numbers <- function(x, y) {
  result <- x * y
  return(result)
}
```

In this example, the multiply_two_numbers function takes two input arguments x and y, multiplies them together, and returns the result.

R functions can also return multiple values using a list. Here's an example of a function that returns multiple values:

```{r}
divide_two_numbers <- function(x, y) {
  quotient <- x / y
  remainder <- x %% y
  return(list(quotient, remainder))
}
```

### Exercises
- Write a function count_words that takes in a string and returns the number of words in the string.
- Write a function mean_sd that takes in a vector of numeric values and returns a list with the mean and standard deviation of the values.



## Vectors
Vectors are an essential data structure in R, and they are used to store collections of values that are of the same type. Vectors in R can be of different types, including numeric, character, logical, complex, and raw. In this article, we will discuss creating vectors, indexing and slicing, and vector operations in R.

### Creating Vectors

To create a vector in R, we use the c() function, which stands for concatenate. We pass values to the c() function, and it returns a vector with those values. Here are some examples:

#### Numeric Vector:
```{r}
# create a numeric vector
num_vec <- c(1, 2, 3, 4, 5)
print(num_vec) # output: [1] 1 2 3 4 5
```

#### Character Vector:
```{r}
# create a character vector
char_vec <- c("apple", "banana", "cherry")
print(char_vec) # output: [1] "apple"  "banana" "cherry"
```

#### Logical Vector
```{r}
# create a logical vector
log_vec <- c(TRUE, FALSE, TRUE)
print(log_vec) # output: [1]  TRUE FALSE  TRUE
```

### Indexing and Slicing
We can access elements of a vector using indexing and slicing. Indexing means accessing a single element of a vector, while slicing means accessing a range of elements. In R, indexing starts at 1. Here are some examples:
```{r}
# create a numeric vector
num_vec <- c(1, 2, 3, 4, 5)

# indexing
print(num_vec[1]) # output: [1] 1

# slicing
print(num_vec[2:4]) # output: [1] 2 3 4
```

### Vector Operations:
We can perform various operations on vectors in R. Here are some examples:

#### Arithmetic Operations:
We can perform arithmetic operations on vectors, such as addition, subtraction, multiplication, and division. In these operations, each element of the vector is operated on individually. Here are some examples:
```{r}
# create two numeric vectors
num_vec1 <- c(1, 2, 3)
num_vec2 <- c(4, 5, 6)

# addition
print(num_vec1 + num_vec2) # output: [1] 5 7 9

# multiplication
print(num_vec1 * num_vec2) # output: [1]  4 10 18

```

#### Logical Operations:
We can perform logical operations on vectors, such as AND, OR, and NOT. In these operations, each element of the vector is operated on individually. Here are some examples:
```{r}
# create two logical vectors
log_vec1 <- c(TRUE, FALSE, TRUE)
log_vec2 <- c(FALSE, TRUE, TRUE)

# AND
print(log_vec1 & log_vec2) # output:
print(log_vec1 | log_vec2)
```

### Exercises
- Create a vector named my_vector containing the numbers 1, 2, 3, 4, 5.
- Create a vector named my_letters containing the letters 'a', 'b', 'c', 'd', 'e'.
- Use the length() function to find the length of my_vector.
- Use the class() function to determine the class of my_letters.
- Use the sum() function to find the sum of my_vector.
- Use the sort() function to sort my_letters in alphabetical order.
- Use the rev() function to reverse the order of my_vector.
- Use the [ ] operator to extract the third element of my_vector.
- Use the [ ] operator to extract the first three elements of my_letters.
- Use the rep() function to create a vector of the number 2 repeated 5 times.

