---
title: "Getting Started With Data Analysis"
author: "Adejumo Ridwan Suleiman"
---

## R Data Types

In R, data types represent the different types of values that can be stored in variables or data structures. The following are the commonly used data types in R.

### Numeric Data Type
Numeric data type is used to represent numbers, including integers and floating-point numbers.
```{r}
# Example:
x <- 10
y <- 3.14
```

### Character Data Type
Character data type is used to represent strings of text.
```{r}
# Example:
name <- "John Doe"

```

### Logical Data Type
Logical data type is used to represent logical values - TRUE or FALSE.
```{r}
# Example:
is_raining <- TRUE
is_sunny <- FALSE

```

### Factor Data Type
Factor data type is used to represent categorical variables. It is useful for handling data with a limited number of possible values.
```{r}
# Example:
gender <- factor(c("male", "female", "male", "male", "female"))

```

### Date and Time Data Types
R has built-in data types for representing dates and times.
```{r}
# Example:
date_today <- Sys.Date()  # returns today's date
time_now <- Sys.time()    # returns the current time

```

### Complex Data Type
Complex data type is used to represent complex numbers.

```{r}
# Example:
z <- 3 + 4i

```


### Raw Data Type
Raw data type is used to represent binary data.
```{r}
# Example:
x <- charToRaw("hello world")
```

### Exercise:

Create a vector that stores the ages of a group of people. The ages should be numeric. Create a factor variable that stores the gender of each person. Finally, create a variable that stores the current date and time.


## R Operators

Operators are used to perform operations on values in R. There are several types of operators in R, including arithmetic operators, assignment operators, comparison operators, logical operators, and more.

### Arithmetic Operators
Arithmetic operators are used to perform mathematical operations on numeric values.
```{r}
# Example:
x <- 10
y <- 3
x + y   # addition
x - y   # subtraction
x * y   # multiplication
x / y   # division
x ^ y   # exponentiation
x %% y  # modulus
x %/% y # integer division
```

### Assignment Operators
Assignment operators are used to assign values to variables.
```{r}
# Example:
x <- 10
y <- 3
x <- y   # assign the value of y to x

```

### Comparison Operators
Comparison operators are used to compare values and return logical values (TRUE or FALSE).
```{r}
# Example:
x <- 10
y <- 3
x == y  # equal to
x != y  # not equal to
x > y   # greater than
x < y   # less than
x >= y  # greater than or equal to
x <= y  # less than or equal to

```


### Logical Operators
Logical operators are used to combine logical values and return logical values.
```{r}
# Example:
x <- TRUE
y <- FALSE
x & y   # AND
x | y   # OR
!x      # NOT

```

### Other Operators
Other Operators
R has other operators, including:

- Membership Operators: `%in%` is used to check if an element is a member of a vector.
- Concatenation Operators: `c()` is used to concatenate vectors.
- Ternary Operator: `ifelse()` is used to return values based on a logical condition.
```{r}
# Example:
x <- 1:5
3 %in% x             # returns TRUE
y <- c(6, 7, 8)
z <- c(x, y)         # concatenates x and y
result <- ifelse(x > 3, "high", "low")  # returns "high" if x > 3, "low" otherwise

```

### Exercise:

Write an R code that creates two variables x and y that store any numeric values of your choice. Using the comparison operators, check if x is greater than y and store the result in a variable called "is_greater". Print out the value of "is_greater".